Evolution of ORM Systems: Formal Foundations

### 2012-12-20 doc/description.tex
updates in removes

\subsubsection{Creation of a System}
\subsubsection{Add Class}

$
\Phi_{addClass}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
}{
\begin{gathered}
  \rho(c) \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ \wedge \rho(associations(c))
\end{gathered}
}
$
\subsubsection{Add Property}
$
\Phi_{addProperty}(p, d) = \begin{cases}
\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
\\ \\ 
\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
 \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \\ \cup pair(name(p), defaultValue(p))
\end{gathered}
} 
\\ \\
\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
 \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), pair("value", defaultValue(p)), \\ pair(name(p), value(primaryKey(r))) 
\end{gathered}
}

\end{cases}
$

\subsubsection{Add Association}
$
\Phi_{AddAssociation}(a, f_m, d) = \begin{cases}
\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), pair(name(a), \\ value(primaryKey(first(r)))) 
\end{gathered}
 }
 \\ \\
\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered}
\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), pair(name(t_s), \\ value(primaryKey(first(r)))) pair(name(t_t), \\ value(primaryKey(second(r)))))
\end{gathered}
 }

 \end{cases}
$

\subsubsection{Deconstruction of a System}
\subsubsection{Remove Class}
$
\Phi_{removeClass}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
$
%TODO remove by label

\subsubsection{Remove Property}
$
\Phi_{removeProperty}(l_c, l_p, d) = \begin{cases}
 \inference{ \exists t \in tables(d) : name(t) = l_c \\ \wedge \exists c \in columns(t) : name(c) = l_p  }{d \vdash
columns(t) \setminus c
} \\ \\
  \inference{ \exists t \in tables(d) : name(t) = l_c \\ \wedge \not \exists c \in columns(t) : name(c) = l_p \\ \wedge \exists u \in tables(d) : name(u) = l_p }{ tables(d) \setminus u }
 \end{cases}
$

\subsubsection{Remove Association}
$
\Phi_{removeAssociation}(l_c, l_a, d) =  \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = l_c 
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_a 
 \\ \wedge not containsData(t)}{ foreignKeys(t) \setminus f }
 \\ \\
 \inference{\exists t \in tables(d) : name(t) = l_a 
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = l_c
 \\ \wedge not containsData(t)}{ tables(d) \setminus t}
 \end{cases}
$

>>>>

\subsubsection{Creation of a System}
\paragraph{Add Class}

$$
\Phi_{addClass}(c, d) = \inference{ \forall t \in tables(d) : name(t) \neq name(\rho(c)) \\
	\forall a \in associations(c), \exists u \in  tables(d) : reference(a) = name(u) 
}{
\begin{gathered}
  \rho(c) \wedge \rho(singleProperties(c)) \wedge \rho(collectionProperties(c)) \\ \wedge \rho(associations(c))
\end{gathered}
}
$$
\paragraph{Add Property}
$$
\Phi_{addProperty}(p, d) = \begin{cases}
\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge (! mandatory(p) \vee ! instantiated(owningClass(p))}{ \rho(p,d)} 
\\ \\ 
\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
 \rho(p,d) \wedge \forall r \in data(d), reference(r) = t : pair(r) \\ \cup pair(name(p), defaultValue(p))
\end{gathered}
} 
\\ \\
\inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge mandatory(p) \wedge instantiated(owningClass(p)) \\ \wedge cardinality(p) \leq 1}{\begin{gathered}
 \rho(p,d) \wedge \forall r \in row(owningClass(p)) : data(d)\\ \cup row(name(p), pair("value", defaultValue(p)), \\ pair(name(p), value(primaryKey(r))) 
\end{gathered}
}

\end{cases}
$$

\paragraph{Add Association}
$$
\Phi_{AddAssociation}(a, f_m, d) = \begin{cases}
\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) \leq 1}{\begin{gathered}
\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)) : \\ data(t_t) \cup row(name(t_s), pair(name(a), \\ value(primaryKey(first(r)))) 
\end{gathered}
 }
 \\ \\
\inference{ \exists t_s, t_t \in tables(d) : name(t_s) = name(owningClass(a)) \\ \wedge name(t_t) = reference(a) \\ \wedge \not \exists f \in foreignKeys(t_t) : name(f) = name(a) \\ startCardinality(a) > 1}{\begin{gathered}
\rho(a,d) \wedge \forall r \in f_m(owningClass(a), reference(a)),\\ t \in tables(d) : name(t) = name(a) : \\ data(t) \cup row(name(a), pair(name(t_s), \\ value(primaryKey(first(r)))) pair(name(t_t), \\ value(primaryKey(second(r)))))
\end{gathered}
 }

 \end{cases}
$$

\subsubsection{Deconstruction of a System}
\paragraph{Remove Class}
$$
\Phi_{removeClass}(c,d) = \inference{ \rho(c) \in tables(d) \\ \wedge \not \exists r \in data(d) : owningTable(r) = \rho(c) \\ \not \exists j \in tables(d), f \in foreignKeys(j) : reference(f) = \rho(c)  }{ tables(d) \setminus \rho(c)}
$$
%TODO remove by label

\paragraph{Remove Property}
$$
\Phi_{removeProperty}(p, d) = \begin{cases}
 \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \exists c \in columns(t) : name(c) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p)) }{columns(t) \setminus c
} \\ \\
  \inference{ \exists t \in tables(d) : name(t) = name(owningClass(p)) \\ \wedge \not \exists c \in columns(t) : name(c) = name(p) \\ \wedge \exists u \in tables(d) : name(u) = name(p) \\ \wedge \not \exists r \in data(d) : owningTable(r) = name(owningClass(p))}{ tables(d) \setminus u }
 \end{cases}
$$

\paragraph{Remove Association}
$$
\Phi_{removeAssociation}(a, d) =  \begin{cases}
 \inference{\exists t \in tables(d) : name(t) = name(owningClass(a))
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = name(a)
 \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ foreignKeys(t) \setminus f }
 \\ \\
 \inference{\exists t \in tables(d) : name(t) = name(a)
 \\ \wedge \exists f \in foreignKeys(t) : name(f) = name(owningClass(a))
 \\ \wedge \forall r \in data(d) : owningTable(r) \neq t}{ tables(d) \setminus t}
 \end{cases}
$$
